---
title: "Functions in R"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a collection of notes about building functions, and some of the best practices that we should emulate when doing so to maintain good code structure. 

## Types of arguments 
There are, generally speaking, two types of arguments:

* __Data arguments__ - _these are the arguments that the result is computed on_
* __Detail arguments__ - _these define how the computation is performed by the function_

Data arguments should always preceed the detail arguments, this helps with clarity, and with the ability to pipe functions together. 


### Argument clarity 
Some older base functions could be renamed and wrapped in a custom user-defined function to aid in the readability and flow of the code. This can be particularly true of model functions that don't take the __data arguments__ as the first argument, and thus can't cleanly be used with pipes and the `tidyverse`. 


Defining our own function could clean up our analysis and code to make it more obvious what we are doing for others to see. Below is an example of this with defining a function that takes the arguments for a `glm()` call and rearranges them, giving a clear name at the same time. 
```{r}
run_poisson_regression <- function(data, formula) {
  glm(formula, data, family = poisson)
}
```

