---
title: "Rowwise functions"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

I can't count the number of times that I've needed to calculate something on a rowwise basis. This probably isn't that common when you have tidy data, and are trying to generate a new metric for every observation in the data, rather than looking at that data one variable at a time.  

I will go through some of the methods that I have used in the past to achieve applying a function per row for whatever reason might have suited my needs at the time.  

- - - -

# Using `apply` 
This may be the most obvious choice but it was one that evaded me for some time as I tried to use `dplyr` and the `tidyverse` to pipe all of my actions together.  

Using `apply` and setting the argument `MARGIN` to 1 means that the function that you use is applied in a rowwise manner. The vector that is then produced as a result can easily be assigned to the data frame that you are working with. 

_Example: you want to work out how many missing values there are per row to find observations with largely missing information._ 

```{r}
df$na_count <- apply(df, MARGIN = 1, ~ sum(is.na(.)))
```

In this case we are going above the usual application and we are defined an anonymous function to achieve what we are looking to achieve. The result from the `apply` call is then assigned to the variable `na_count` on the data frame `df`; you could later filter by this to give you access to just those observations with a certain number of missing values, or above a threshold that you have defined. 

- - - -

# Using `do` 
Another method that I have used in the past, mainly due to the fact that you can incorporate it into pipelines (using `%>%`) is by calling the function `do` on a row by row basis.  

I would probably wager that this isn't as efficient as the `apply` method, but nonetheless it is an option if you are looking to apply a method by a row. It probably offers slightly more capability compared to the apply, as you can do as much computation within the section as you like. 

Note that I also use the `dplyr` function `rowwise()` here to 

_Following the same example as before_

```{r eval = TRUE}
library(tidyverse)
df <- expand.grid(x = 1:3, y = 3:1, z = NA)
df %>%
  rowwise() %>%
  do({
    # Convert the row into a tibble
    row <- as_tibble(.)
    # Mutate a new column for the row
    row <- row %>% mutate(na_count = sum(is.na(.)))
    # Return the altered row
    row
  })
```  

This is obviously a slightly over the top example, but I am just highlighting a way in which you can achieve a function, or possibly set of functions, being called by row. 






