---
title: "Betting Simulations"
output: rmarkdown::github_document
---

```{r define_global_settings, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, error=FALSE, message=FALSE)
```

Load the relevant packages that I will be using during these simple simulations. 

```{r load_packages}
library(tidyverse)
library(reshape2)
```

There are a lot of systems out there for trying to make betting a consistent and reliable process. Something that is, inherently, neither of those things. 

Below I define a simulator function that defines a number of basic parameters that can start you off in simulating how betting methods would work if applied over a number of individual bets. 

```{r define_simulator_function}
simulator <- function(starting_bank = 10, 
                      n_bet = 20, 
                      stake_proportion = 0.25, 
                      odds_range = c(1.6, 1.8), 
                      win_perc = 50){
  # Loop through n_bet and simulate the result
  bank = starting_bank;
  bank_tracker = c(bank)
  
  for(i in 1:n_bet) {
    # Pick a random odd between the range defined
    odds <- sample(seq(odds_range[1], odds_range[2], 0.01), 1)
    chance <- sample(1:100, 1)
    
    # Did they win that particular bet? 
    if(chance <= win_perc){
      bank = round((bank * (1 - stake_proportion)) + (bank * stake_proportion * odds), 2)
      stake <- bank * stake_proportion
      winnings <- (bank * stake_proportion * odds)
      print(paste("Won bet ", i, ": ", bank, sep = "" ))
    } else {
      bank = round(bank * (1 - stake_proportion), 2)
      print(paste("Lost bet ", i, ": ", bank, sep = "" ))
    }
    bank_tracker = c(bank_tracker,  bank)
  }
  return(bank_tracker)
}
```

*starting_bank*: the amount of money that you are starting with.  
*n_bet*: the number of bets you would like to simulate.  
*stake_proportion*: the proportion of your bank you're staking on each bet (constant).  
*odds_range*: the range of odds that you're betting on.  
*win_perc*: the average win percentage within the odds range.  

```{r generate_single_simulation}
set.seed(100)
(results <- simulator())
```

```{r single_simulation_plotting, fig.dim=c(10,4)}
tibble(bank = results, 
       n_bet = 1:length(results), 
       in_profit = ifelse(bank >= 10, T, F)) %>%
  ggplot(aes(n_bet, bank)) + 
    geom_hline(yintercept = results[1], alpha = 0.25) + 
    geom_line(color = '#7f8c8d') + 
    geom_point(aes(color = in_profit)) + 
    theme_minimal() + 
    scale_colour_manual(values = c("TRUE" = "#27ae60", "FALSE" = "#e74c3c")) +
    labs(title = "Single simulation plotted")
```

A good start to the simulation meant that we were in profit. The longer the betting goes on, you would think that the more likely you are to be in the negative. 