---
title: "Betting Simulations"
output: rmarkdown::github_document
---

```{r include=FALSE}
knitr::opts_chunk$set(warning=FALSE, error=FALSE, message=FALSE)
```

Load the relevant packages that I will be using during these simple simulations. 

```{r}
library(tidyverse)
library(reshape2)
```

There are a lot of systems out there for trying to make betting a consistent and reliable process. Something that is, inherently, neither of those things. 

Below I define a simulator function that defines a number of basic parameters that can start you off in simulating how betting methods would work if applied over a number of individual bets. 

```{r}
simulator <- function(starting_bank = 10, 
                      n_bet = 20, 
                      stake_proportion = 0.25, 
                      odds_range = c(1.6, 1.8), 
                      win_perc = 50){
  # Loop through n_bet and simulate the result
  bank = starting_bank;
  bank_tracker = c(bank)
  
  for(i in 1:n_bet) {
    # Pick a random odd between the range defined
    odds <- sample(seq(odds_range[1], odds_range[2], 0.01), 1)
    chance <- sample(1:100, 1)
    
    # Did they win that particular bet? 
    if(chance <= win_perc){
      bank = round((bank * (1 - stake_proportion)) + (bank * stake_proportion * odds), 2)
      stake <- bank * stake_proportion
      winnings <- (bank * stake_proportion * odds)
      print(paste("Won bet ", i, ": ", bank, sep = "" ))
    } else {
      bank = round(bank * (1 - stake_proportion), 2)
      print(paste("Lost bet ", i, ": ", bank, sep = "" ))
    }
    bank_tracker = c(bank_tracker,  bank)
  }
  return(bank_tracker)
}
```

*starting_bank*: the amount of money that you are starting with.  
*n_bet*: the number of bets you would like to simulate.  
*stake_proportion*: the proportion of your bank you're staking on each bet (constant).  
*odds_range*: the range of odds that you're betting on.  
*win_perc*: the average win percentage within the odds range.  

```{r}
set.seed(100)
simulator()
```

