---
title: "Building and comparing models with `caret`"
output: 
  rmarkdown::github_document:
    toc: true
---

`caret` is an immensely powerful [package](http://topepo.github.io/caret/index.html) for R which allows for the control of training parameters, reproducible model comparison, quick model comparison, tuning of hyperparameters and preprocessing of data, among others.  

I've had a fair bit of experience building models with it, and trying to learn the best practices of machine learning and model building. Those projects can primarily be found on my [Kaggle profile](https://www.kaggle.com/willcanniford), or there is some record of them here on [my github](https://github.com/willcanniford/kaggle).

## Small worked example using the `Sonar` dataset
The `Sonar` dataset is a classic example dataset that can be obtained through the `mlbench` package. I'll load and just show the first few rows to get an idea for the structure and what we might be able to do with it.  

```{r load_package_and_data, message=FALSE}
library(tidyverse)
library(mlbench);data(Sonar)
glimpse(Sonar)
```
  
We've got 60 numeric variables, and then a single factor variable named `Class`. Let's just have a look and see how many levels the factor has, as that will alter the way that we approach the problem, and what algorithms we use.  

_*Note*_: this is a *classification* problem as we are predicting a category. 

```{r explore_response_levels}
glimpse(Sonar$Class)
```  

Ok, we've got 2 classes, so we can start with a logistic regression, but we also have algorithms like knn, random forest and support vector machine that we can also use. 
